Project 1 Checkpoint

### 2015-02-27 project1/checkpoint/main.tex
Basic edits in preparation for the final report

\section{The Conceptual Model}
In order to create a representative software simulation of the aforementioned system, we first begin by designing a
conceptual version of this model from which we can plan our implementation. We'll borrow heavily from the results and data
in \cite{cts12} when formulating our model.
Our architecture will be event driven, prioritizing events by timestamp. Furthermore,
we'll utilize the publisher-subscriber pattern (see \cite{pubsub}) to promote better decoupling.
Given this architecture, our first step will be to identify the entities present in our model, and then consider their
interactions (or the inherent ``activies'') within the simulation. This will allow us then to design the events that
are important

\subsection{Entities}
The first entity, unsurprisingly, is a single car. Our representation of a car is clearly less detailed than the real
thing, but rather focuses on variables that influence it's travel time from its starting location to destination, along
with the variables that influence neighboring cars. For our simulation, we'll track the position and destination of the
cars, along with the size of the car since that can influence the position of cars behind it. Also, we will track the
current velocity of the car

@@

An intersection serves as holding point for cars traveling along the road so if a light is red, cars are forced to
queue up at the intersection until green. This lends itself to a queue-style data structure that encorporates both
position and acceleration. Car acceleration is a complex topic, but models have been constructed \cite{bonneson}
\cite{herman_et_al}. These models are generally continuous, and can generally be represented as a differential equation
due to the relationships between acceleration and velocity, as can be seen in \cite{briggs} \cite{deceleration} as well.
Instead, we will
simply consider the various events that trigger changes in behavior. For instance, if a car slows down, the car behind
it will be notified and change it's driving patterns accordingly.

@@

We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
consider basic aspects of a car: when it enters the system, and it's intended destination (turning left or right at
12th street or going straight). We will assume that all cars have the same speed entering the system, and that they
have the same length. These values can be calculated from the NGSIM data set \cite{ngsim}.

@@

The distribution of car arrival is very important, but has been difficult to model. The arrival intervals between 
northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
fits. Furthermore, pruning outliers via the common interquartile range approach, where values are rejected when
they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.

@@

\subsection{Simulation Output}
The most important statistic we'll want to collect during the simulation is the thoughput of the simulation. We'll
want to observe how the light timing affects the average travel time for cars. It will also to see how variable this
is. For instance, do cars travelling on Peachtree St. have similar travel times, despite their arrival time, or does
their arrival drastically lengthen their time to travel the stretch of road we are modeling.

@@

Another simplification to consider is lane changing. Lane changing models do exist \cite{lanechanging} \cite{lanechanging2},
but can be complex, requiring modeling of driver choices and gap acceptance for drivers. Peachtree is a two lane road,
so lane changes are possible. However, we will make the assumption that drivers by this point have already made any
discretionary lane changes and traffic is evened out between lanes. Also, we will assume left turning drivers will be
in the left lane already. Similarly, right turning drivers will already be in the right lane.

>>>>

\section{The Conceptual Model}
In order to create a representative software simulation of the aforementioned system, we first begin by designing a
conceptual version of this model from which we can plan our implementation. We'll borrow heavily from the results and data
in \cite{cts12} when formulating our model.
Our architecture will be event driven, prioritizing events by timestamp. Furthermore,
we'll utilize the publisher-subscriber pattern (see \cite{pubsub}) to promote better decoupling.
Given this architecture, our first step will be to identify the entities present in our model, and then consider their
interactions (or the inherent ``activies'') within the simulation. This will allow us then to design the events that
are important and relevant to the simulation.

\subsection{Entities}
The first entity, unsurprisingly, is a single car. Our representation of a car is clearly less detailed than the real
thing, but rather focuses on variables that influence it's travel time from its starting location to destination, along
with the variables that influence neighboring cars. For our simulation, we'll track the position and destination of the
cars. Probably the most primary value associated with a car is its arrival time, since that dictates when and where it
will fall in the queue of cars at the intersection.

@@

An intersection serves as holding point for cars traveling along the road so if a light is red, cars are forced to
queue up at the intersection until green. This lends itself to a queue-style data structure that encorporates both
position and acceleration. Car acceleration is a complex topic, but models have been constructed \cite{bonneson}
\cite{herman_et_al}. These models are generally continuous, and can generally be represented as a differential equation
due to the relationships between acceleration and velocity, as can be seen in \cite{briggs} \cite{deceleration} as well.
Instead, we will simply consider the various events that trigger changes in behavior. For instance, a real driver is
constantly observing their environment (hopefully) and making minute adjustments the entire time. However, we'll take
a greatly simplified approach and simply precalculate an estimate of how long a car takes to traverse the section of
road, along with a model for accelerating from a stop.

@@

We first concern ourselves with finding an input distribution for cars entering the system. In our model, we'll only
consider basic aspects of a car: when it enters the system, and it's intended destination (north or south). We will assume that all cars have the same speed entering the system, and that they
have the same length. These values can be calculated from the NGSIM data set \cite{ngsim}.

@@

The distribution of car arrival is very important, but has been difficult to model. The arrival intervals between 
northbound and southbound cars are provided in \ref{fig:northbound_arrival} and \ref{fig:southbound_arrival}.
Model fitting was attempted for both $\Gamma$ and $\Gamma$-mixture models on the raw data, but resulted in poor
fits. Furthermore, pruning outliers was attempted as well, but still yielded bad matches. The outlier critera used for
this determination is the common interquartile range approach, where values are rejected when
they lie outside of $median + 1.5IQR$, where $IQR$ is the interquartile range $Q_3 - Q_1$.

@@

\subsection{Simulation Output}
The most important statistic we'll want to collect during the simulation is the thoughput of the simulation. We'll
want to observe how the light timing affects the average travel time for cars. It will also to see how variable this
is. For instance, do cars travelling on Peachtree St. have similar travel times, despite their arrival time, or does
their arrival drastically lengthen their time to travel the stretch of road we are modeling. Furthermore, we might also
ask if the light timing has that much of an effect at all, and if it does what is the ``critical point`` at which the
system no longer can serve incoming cars fast enough?

@@

Another simplification to consider is lane changing. Lane changing models do exist \cite{lanechanging} \cite{lanechanging2},
but can be complex, requiring modeling of driver choices and gap acceptance for drivers. Peachtree is a two lane road,
so lane changes are possible. However, we will make the assumption that drivers by this point have already made any
discretionary lane changes and traffic is evened out between lanes. This allows us to consider only a single lane at a
time, since the second lane is identical to the first after these assumptions.

We also assume here that cars are entering at the same velocity, based off of the urban speed limit of 30 mph set by
the state of Georgia (CITE). Furthermore, we assume a single sized car, based off of the average from the NGSIM data.
Having a constant speed and length allows us to simplify the calculation of stop points for cars entering the
intersection. Although these simplifications help, dealing with acceleration is an even more complex matter. In this
case it is both a mixture of both discrete (the light signal) and continuous elements (the car's velocity). By choosing
an event-driven approach we are forced to simplify acceleration, since we cannot accurately represent a continuous
system in this manner.

Due to our inability to map a common distribution to the data gathered in the NGSIM study, we are using an empircal
distribution to model the input distribution of cars. Although this is an "easy" choice, it comes with the assumption
that both the measured data is representative of the true input distribution of cars, and that there is no measurement
error in the data collected.
